---
output:
  md_document: 
    preserve_yaml: false
  html_document:
    keep_md: yes 
---

# PSEM: Proxy Spectral Error Model.

-------


```{r knitr_setup, echo=FALSE}
knitr::opts_chunk$set(dev = "svg")
```

PSEM implements the Proxy Spectral Error Model described in the discussion papers:

* Estimating the timescale-dependent uncertainty of paleoclimate records—a spectral approach. Part I: Theoretical concept 
* Estimating the timescale-dependent uncertainty of paleoclimate records—a spectral approach. Part II: Application and interpretation.


## Installation 

**psem** can be installed directly from github

```{r, eval=FALSE}
if (!require("remotes")) {
  install.packages("remotes")
}

remotes::install_github("EarthSystemDiagnostics/psem")
```


## Usage

```{r}
library(psem)
```


### Parametrise a proxy error spectrum for a core at 45°N 0°E


#### Power spectrum for the stochastic climate

```{r climate_spec_ex1, fig.width=4, fig.height=3}
# PSD Climate
example.lat <- 45

clim.spec.ex1 <- ModelSpectrum(
  freq = NULL,
  latitude = example.lat,
  variable = "temperature", beta = 1
)

p.clim.spec.ex1 <- PlotModelSpectrum(clim.spec.ex1)
p.clim.spec.ex1
```


#### Amplitude of the seasonal cycle

```{r}
seasonal.amp <- AmpFromLocation(
  longitude = 0,
  latitude = example.lat,
  proxy.type = "degC",
  depth.upr = 0, depth.lwr = -50
)
```

#### Orbital modulation of the amplitude of the seasonal cycle
```{r}
orbital.pars <- RelativeAmplitudeModulation(
  latitude = example.lat,
  maxTimeKYear = 23,
  minTimeKYear = 1,
  bPlot = FALSE
)
```

#### Get list of parameters

```{r}
# sediment accumulation rate for the core
ex.sed.acc.rate <- 10

spec.pars.ex1 <- GetSpecPars(
  proxy.type = "Mg_Ca",
  T = 1e04,
  delta_t = 100,
  tau_r = 100,
  sig.sq_a = orbital.pars$sig.sq_a,
  sig.sq_c = seasonal.amp$sig.sq_c,
  tau_b = 1000 * 10 / ex.sed.acc.rate,
  tau_s = 1000 * 1 / ex.sed.acc.rate,
  N = 30,
  tau_p = 7/12,
  phi_c = 0, delta_phi_c = 2 * pi / 3,
  phi_a = pi / 2,
  sigma.cal = 0.25,
  sigma.meas = 0.25,
  sigma.ind = 1,
  clim.spec.fun = "ModelSpectrum",
  clim.spec.fun.args =
    list(latitude = example.lat, beta = 1)
)


```

#### Call `ProxyErrorSpectrum` with these parameters and plot it.

```{r, fig.width=5, fig.height=3}
proxy.err.spec <- do.call(ProxyErrorSpectrum, spec.pars.ex1)
PlotSpecError(proxy.err.spec)
```

#### Integrate the error spectrum to get timescale-dependent error

```{r, fig.width=4.5, fig.height=3}
tsd.error.var <- IntegrateErrorSpectra(proxy.err.spec)
PlotTSDVariance(tsd.error.var)
```

#### Get error for a record smoothed to a given timescale

```{r}
err.500 <- GetProxyError(tsd.error.var, timescale = 500)
knitr::kable(err.500, digits = 2)
```

